Psychophysics of musical elements in the discrete-time representation of sound

### 2013-10-07 article/article.tex
feito ateh arcos ciclicos

The arcs can be decomposed in two convergent sequences: the first arc reaches the
apex and the second arc returns from the apex to the start region. This apex is called climax by traditional music theory. It is
possible to differ between arcs whose climax are placed at the beginning, middle, end,
and at the first and second half of the considered duration. These structures are
shown in Figure~\ref{fig:climax}. The parameter of variation does not exist, thus
the arc consists only of the reference structure~\cite{Schoenberg}.

\begin{figure}
    \centering
        \includegraphics[width=\columnwidth]{figures/climax}
        \caption{Canonical distinctions of musical climax in a given melody and
        other domains. The considered possibilities are: climax at beginning,
        climax at first half, climax at middle, climax at second half and climax
        at the end. The y-axis is not properly specified since there exist no
        real parametric variation and, therefore, the structure is a reference.}
        \label{fig:climax}
\end{figure}

Consider $S_i=\{s_i\}_0^{H-1}$ as an increasing sequence. The sequence
$R_i=\{r_i\}_0^{2H -2}=\left\{s_{(H-1-|H-1-i|)}\right\}_0^{2H-2}$ is a sequence
that presents a perfect profiteer symmetry, i.e.\ the second half is the
mirrored version of the former. Following the musical concepts, the climax is
localized in the middle of the sequence. It is possible to modify this
fact using sequences with different sizes. All the mathematical theory of
sequences, already well established and taught routinely in calculus courses, can be
used to generate those arcs~\cite{Guidorizzo,Schoenberg}. 
Theoretically, when applied to any characteristic of music events, these sequences produce arcs, since they imply the deviation and return of an initial
parametrization. Henceforth, it is possible to have a
number of distinct arcs, with different sizes and climax for a same sequence of events. This is an interesting and useful resource, and the correlation of arcs results in coherence~\cite{Salzer}.

Historically (and practically nowadays), the golden ratio and the Fibonacci sequence have special importance. The Lucas sequence allows the generalization of
Fibonacci sequence, making it easy to understand. Given any two numbers $x_0$
and $x_1$, the Lucas sequence can be obtained as: $x_n=x_{n-1}+x_{n-2}$. The greater $n$ is, the more $\frac{x_{n}}{x_{n_1}}$ approaches the golden ratio:
$1.61803398875...$. The sequence converges fast even with high discrepant
initial values. Being $x_0=1$ and $x_1=100$, and $y_n=\frac{x_n}{x_{n+1}}$ an
auxiliary sequence, the error occurring in the first values with
relation to the golden ratio is, approximately, $\{ e_n \}
=\left\{100\frac{y_n}{1.61803398875}-100 \right\}_1^{10}=\{6080.33, -37.57, 23,
-7.14, 2.937, -1.09, 0.42, -0.1601, \\ 0.06125, -0.02338\}$. The Fibonacci sequence
presents about the same error progression but starts right at the second step since $\frac{1}{1}\approx\frac{101}{100}$.

The music piece \emph{Dirracional} exposes the use of arcs into directional structures.
Its source code is included in Appendix~\ref{ap:dirracional} and available online as part
of \massa~\cite{MASSA}.

>>>>

The arcs can be decomposed in two convergent sequences: the first reaches the
apex and the second returns from apex to start region.
This apex is called climax by traditional music theory. It is
possible to differ between arcs whose climax are placed at the beginning, middle, end,
and at the first and second half of the considered duration. These structures are
shown in figure~\ref{fig:climax}. The varying parameter can be unexistent, a case in which
the arc consists only of a reference structure~\cite{Schoenberg}.

\begin{figure}
    \centering
        \includegraphics[width=\columnwidth]{figures/climax}
        \caption{Canonical distinctions of musical climax in a given melody and
        other domains. Considered possibilities are: climax at beginning,
        climax at first half, climax at middle, climax at second half and climax
        at the end. The x and y-axis is not properly specified since the parameters
        can be non-existent, as in a reference structure.}
        \label{fig:climax}
\end{figure}

Consider the $S_i=\{s_i\}_0^{H-1}$ increasing sequence. The sequence
$R_i=\{r_i\}_0^{2H -2}=\left\{s_{(H-1-|H-1-i|)}\right\}_0^{2H-2}$
presents perfect specular symmetry, i.e.\ the second half is the
mirrored version of the first. In musical terms, the climax is
in the middle of the sequence. It is possible to modify this
by using sequences with different sizes. All the mathematics of
sequences, already well established and taught routinely in calculus courses, can be
used to generate these arcs~\cite{Guidorizzo,Schoenberg}. 
Theoretically, when applied to any characteristic of musical events, 
these sequences produce arcs, since they imply a deviation and return of an initial
parametrization.
Henceforth, it is possible for an only sequence to have
numberous distinct arcs, with different sizes and climax. 
This is an interesting and useful resource, and the correlation of arcs yields coherence~\cite{Salzer}.

In practice, and historically, there is special incidence and use of the golden ratio.
The Lucas sequence allows the generalization of
Fibonacci sequence, making it easy to follow. Given any two numbers $x_0$
and $x_1$, the Lucas sequence can be obtained as: $x_n=x_{n-1}+x_{n-2}$. The greater $n$ is, the more $\frac{x_{n}}{x_{n_1}}$ approaches the golden ratio:
$1.61803398875...$. The sequence converges fast even with high discrepant
initial values. Being $x_0=1$ and $x_1=100$, and $y_n=\frac{x_n}{x_{n+1}}$ an
auxiliary sequence, the error for the fraction of first values with
respect to the golden ratio is, approximately, $\{ e_n \}
=\left\{100\frac{y_n}{1.61803398875}-100 \right\}_1^{10}=\{6080.33, -37.57, 23,
-7.14, 2.937, -1.09, 0.42, -0.1601, \\ 0.06125, -0.02338\}$. The Fibonacci sequence
presents exactly the same error progression, but starts at the second step of the most discrepant
case ($\frac{1}{1}\approx\frac{100+1 = 101}{100}$).

The music piece \emph{Dirracional} exposes the use of arcs into directional structures.
Its source code is available online as part
of \massa\ toolbox~\cite{MASSA}.
